Layered Formal Verification of an Industrial TCP/IP Stack

### 2021-06-01 SecDev21/paper.tex
Rephrasing the se section

In the synchronous scenario of interaction between the TCP tasks, the resulting state of a user function can be affected by the rest of the TCP tasks when the user's functions release the mutex to allow for other events to occur. In this case, the function \spark{TCP_Wait_For_Events} is called from the user function to check if the event requested is completed. For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely the \spark{SOCKET\_EVENT\_CONNECTED}. The call to the \spark{TCP_Wait_For_Events} releases the mutex, and when the \emph{arriving} task receives the appropriate segments, in this case, a segment with the \syn flag, the connection will be established. This will move the socket to the \sestab state. In the meantime, the \spark{TCP\_Update\_Events} function is monitoring for any state changes, and when it notices that the connection is established, it will trigger the \spark{SOCKET\_EVENT\_CONNECTED} event, using the OS event mechanism. This will allow the \spark{TCP\_Connect} user function to resume execution.

@@

\todo[inline]{KG to GC : when you are talking about functions you should say where they are located. There will be readers that will want to see the corresponding code in the repository. Look for examples in the HICLASS report: e.g. This is implemented by the Tcp\_Wait\_For\_Events\_Proof function, located in the tcp\_misc\_binding.adb}

@@

To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} is introduced.
The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} shown in \Cref{algo:waitForEvents}. It computes the set of possible final states after the completion of an event by reusing the function \spark{TCP_Process_One_Segment}. Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}. This ghost function is proved automatically without even a loop invariant, thanks to loop unrolling in GNATprove.

\todo[inline]{GC: modified up to here}

@@

The first step in order to extract contracts was to read the state machine to
know all the possible transitions that can be done by the reception of a flag,
and adapt it to
the subtleties of CycloneTCP. The next step is to write a driver to run KLEE and
know if the contracts that we have inferred are correct or not.
A simplified example is shown in \Cref{code:kleedriver}. The first lines
create a symbolic incoming segment to process and a symbolic socket.
At \cref{code:kleedriver:tcpProcessSegment} the function that processes the
segment is executed. It is followed at \cref{code:kleedriver:assert} by a KLEE
builtin that checks if the code leads to a valid state identified in the
protocol’s specification for all the paths explored.

Once the contract is validated by KLEE, it can be copied out on the SPARK
function \spark{TCP_Process_One_Segment}. With this step done, the concurrency
challenge has been solved and the function
\spark{TCP_Wait_For_Events} has a correct and an accurate contract, which
makes possible the verification of the functional correctness of the TCP user
functions thanks to SPARK.

@@

However weeknesses exist in our implementation, in particular because all the
underlying layers are still written in C. The principal function of those layers
is to format packet before they are sent, or parse incoming packets, check their
integrity and transmit the resulting payload to the corresponding upper level
layer. This processing part can be a source of errors and bugs as explain
in~\cite{Reiher2019RecordFluxFM}.
Using the RecordFlux DSL to parse the packets of the different protocols is the
next step to make the stack safer. Using RecordFlux to parse TCP message
would be a big step to finish the translation of the TCP protocol in SPARK.
The postconditions extracted by KLEE could be proved by GNATprove for a
safer result.

>>>>

In the synchronous scenario of interaction between the TCP tasks, the resulting state of a user function can be affected by the rest of the TCP tasks when the user's functions release the mutex to allow for other events to occur. In this case, the function \spark{TCP_Wait_For_Events} (located in the file \texttt{tcp\_misc\_binding.ads}) is called from the user function to check if the event requested is completed. For example, the \spark{TCP_Connect} user function calls \spark{TCP_Wait_For_Events} to wait for the completion of the socket connection event, namely the \spark{SOCKET\_EVENT\_CONNECTED}. The call to the \spark{TCP_Wait_For_Events} releases the mutex, and when the \emph{arriving} task receives the appropriate segments, in this case, a segment with the \syn flag, the connection will be established. This will move the socket to the \sestab state. In the meantime, the \spark{TCP\_Update\_Events} function is monitoring for any state changes, and when it notices that the connection is established, it will trigger the \spark{SOCKET\_EVENT\_CONNECTED} event, using the OS event mechanism. This will allow the \spark{TCP\_Connect} user function to resume execution.

@@

% \todo[inline]{KG to GC : when you are talking about functions you should say where they are located. There will be readers that will want to see the corresponding code in the repository. Look for examples in the HICLASS report: e.g. This is implemented by the Tcp\_Wait\_For\_Events\_Proof function, located in the tcp\_misc\_binding.adb}

@@

To model any possible state changes upon the reception of a segment in SPARK, the function \spark{TCP_Process_One_Segment} located in the file \texttt{tcp\_fsm\_binding.ads} is introduced.
The behavior of \spark{TCP_Wait_For_Events} is modeled by the ghost function \spark{TCP_Wait_For_Events_Proof} (that can be found in the file \texttt{tcp\_misc\_binding.ads}) shown in \Cref{algo:waitForEvents}. It computes the set of possible final states after the completion of an event by reusing the function \spark{TCP_Process_One_Segment}. Both functions have the same behavior, and thus, the contract of the function \spark{TCP_Wait_For_Events_Proof} is also attached to \spark{TCP_Wait_For_Events}. This ghost function is proved automatically without even a loop invariant, thanks to loop unrolling in GNATprove.

@@

We want to prove that the C code implementation of the respects rules extracted from RFC 793. To achieve this, symbolic execution is used to verify that all the possible paths when executing the \spark{tcpProcessSegment} function will result in a state that respects the RFC 793 rules. This process requires three steps, that are show in \Cref{code:kleedriver}. A symbolic incoming segment is created in the first lines to allow us to account for all possible incoming segments. Then at \cref{code:kleedriver:tcpProcessSegment} the function \texttt{tcpProcessSegment} is executed symbolically with KLEE. Finally, it is followed at \cref{code:kleedriver:assert} by a call to \texttt{klee\_assert} that checks if the code leads to a valid state identified in the protocol’s specification.

Once the \texttt{klee\_assert} is validated by KLEE, which means that the code respects the RFC 793 specification, it can be copied out as a valid SPARK contract for the function \spark{TCP_Process_One_Segment}. With this step done, the concurrency challenge has been solved and the function \spark{TCP_Wait_For_Events} has a correct and an accurate contract, which makes possible the verification of the functional correctness of the TCP user
functions thanks to SPARK.

@@

However weaknesses exist in our implementation, in particular because all the
underlying layers are still written in C. The principal function of those layers
is to format packet before they are sent, or parse incoming packets, check their
integrity and transmit the resulting payload to the corresponding upper level
layer. This processing part can be a source of errors and bugs as explain
in~\cite{Reiher2019RecordFluxFM}.
Using the RecordFlux DSL to parse the packets of the different protocols is the
next step to make the stack safer. Using RecordFlux to parse TCP message
would be a big step to finish the translation of the TCP protocol in SPARK.
The postconditions extracted by KLEE could be proved by GNATprove for a
safer result.
